The Part-Time Parliament\hfill\insertframenumber/\inserttotalframenumber

### 2011-05-25 disciplinas/usi_algsem/paxosTalk/paxosTalk.tex
26 slides, 59 pages

@@

    \vspace{4 mm}
    \item $B2(\beta)$ %B2 may be ensured by always picking a majority of processes as quorum, which implies that n > 2f (a majority is correct for each instance)

@@

    \vspace{4 mm}
    \item $B3(\beta)$

@@

\vspace{3 mm}
\textsc{\textbf{Proof of Lemma}}. Let $\Psi(B, \beta)$ be defined as: $$\Psi(B, \beta) = \{B' \in \beta : (B'_{bal} > \bbalm) \wedge (B'_{val} \neq \bvalm)\}$$

@@

\vspace{3 mm}
(1) As $\Psi(B, \beta) \neq \emptyset$, we pick $C \in \Psi(B, \beta)$, such that $C_{bal} = min\{B'_{bal} : B' \in \Psi(B, \beta)\}$

\vspace{2 mm}
(2) From (1) and from the definition of $\Psi(B, \beta)$, we have that $C_{bal} > \bbalm$

@@

\vspace{2 mm}
(4) As $C_{bal} > \bbalm$ and $\bvotm \cap C_{qrm} \neq \emptyset$, we have that \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)_{bal} \geq \bbalm$

\vspace{2 mm}
(5) From (4) and the definition of \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)$, we know that such vote exists in \textit{Votes}$(\beta)$ (i.e. it is not $null$).

\vspace{2 mm}
(6) From (5) and $B3(\beta)$, we have that \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)_{val} = C_{val}$

\vspace{2 mm}
(7) From (6), and as $C_{val} \neq \bvalm$ (from $\Psi(B, \beta$)), we have that \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)_{val} \neq \bvalm$

@@

\vspace{2 mm}
(9) From (8), and since \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)_{val} \neq \bvalm$, we have that \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta) \in \text{\textit{Votes}}(\Psi(B,\beta))$

\vspace{2 mm}
(10) By the definition of \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)$, we have that \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)_{bal} < C_{bal}$

\vspace{2 mm}
(11) From (10), we have that $\exists B' : B' \in \Psi(B, \beta) \wedge B'_{bal} < C_{bal}$, which means that $$C_{bal} \neq min\{B'_{bal} : B' \in \Psi(B, \beta)\}$$ which is a contradiction with (1). $\square$

@@

\vspace{2 mm}
\textbf{\textsc{Proof of Theorem}}. If $B'_{bal} = B_{bal}$ then $B1(\beta)$ implies $B' = B$. If $B'_{bal} \neq B_{bal}$, the theorem follows immediately from the lemma. $\square$

@@

\vspace{2 mm}
\textbf{\textsc{Proof of Theorem}}.
\begin{itemize}
\item $B1(\beta \cup \{B'\})$ follows from $B1(\beta)$ and the fact that $B'_{bal} = b$, and the assumption about $b$.
\item $B2(\beta \cup \{B'\})$ follows from $B2(\beta)$ and the fact that $B'_{qrm} = Q$, and the assumption about $Q$.
\item For $B3(\beta \cup \{B'\})$, if \textit{MaxVote}$(b,Q,B) = -\infty$ then let $B'_{val}$ be any value; otherwise, let $B'_{val} = \text{\textit{MaxVote}}(b,Q,B)_{val}$. $B3(\beta \cup \{B'\})$ then follows from $B3(\beta)$. $\square$
\end{itemize} 
\end{frame}

%==========================================================
\section{The Paxos Algorithm}
%==========================================================

@@

\begin{itemize}
  \item To maintain $B1(\beta)$, whenever $p$ starts a new ballot $B$, it assigns a sequence number $seq$, which was never used by $p$. To avoid collisions, $p$'s unique id is appended. Then $\bbalm = (seq, p_{id})$.
  \item To maintain $B2(\beta)$, each ballot $B$ has \bqrm\ as the majority set of the processes. As it is impossible to choose two disjoint majority sets from the same group*, $B2(\beta)$ is satisfied.
  \item To maintain $B3(\beta)$, before initiating a ballot $B$, $p$ has to find out what is \textit{MaxVote}$(B, \bqrmm, \beta)$, which is done by finding the \textit{MaxVote}$(B, q, \beta)$ for each $q$ in \bqrm.
  \begin{itemize}
    \item If \textit{MaxVote}$(B, \bqrmm, \beta) = null$, then \bval\ can be anything;\\
    otherwise $\bvalm = \text{\textit{MaxVote}}(B, \bqrmm, \beta)_{val}$.
  \end{itemize}
\end{itemize}

{\small * If half of the processes is faulty, consistency is still respected and there is no deadlock, but the ballot $B$ does not succeed}

@@

  \hspace{3 mm} \textbf{Phase 1a}) $p$ picks a ballot number $b$ and sends \textit{NextBallot}$(b)$ to other processes

  \hspace{3 mm} \textbf{Phase 1b}) each process $q$ responds with \textit{LastVote}$(b, v)$ to $p$, where $v$ is the vote cast by $q$ with the highest ballot number less than $b$ (or $null$)

@@

  Phase 2) Execute the ballot $B$

  \hspace{3 mm} \textbf{Phase 2a}) after receiving \textit{LastVote}$(b, v)$ from a majority $Q$ of the processes, $p$ initiates the ballot with number $b$ and value $val$ such that $B3(\beta)$ is maintained, by sending \textit{BeginBallot}$(b, val)$ to every process in $Q$%Bqrm = Q, Bval = val, Bbal = b

  \hspace{3 mm} \textbf{Phase 2b}) upon receiving \textit{BeginBallot}$(b, val)$, $q$ will not vote on it if it has already sent \textit{LastVote}$(b', v')$, such that $b' > b$; otherwise, it sends \textit{Voted}$(b, q)$ to $p$ and records this vote in its memory

@@

  \vspace{6mm}
  Every phase of this protocol is optional. Consistency is still maintained, although progress is not guaranteed.

@@

  Phase 2)

@@

  \vspace{6mm}
  Consistency is still maintained, and progress is still not guaranteed.

@@

%----------------------------------------------------------
\subsection{Complete (synchronous) protocol}
%----------------------------------------------------------

\begin{frame}
\frametitle{Complete (synchronous) protocol}

@@

%==========================================================
\section*{Summary}
%==========================================================

\begin{frame}
\frametitle<presentation>{Summary}

\begin{itemize}
  \item The \alert{first main message} of your talk in one or two lines.
\end{itemize}

% The following outlook is optional.
\vskip0pt plus.5fill
\begin{itemize}
  \item Outlook
  \begin{itemize}
    \item Something you haven't solved.
    \item Something else you haven't solved.
  \end{itemize}
\end{itemize}

@@

>>>>

  \vspace{3mm}

  \pause

  Some people do use the described \textbf{Paxos} protocol: % 20 years after being first described?...

  \begin{itemize}
    \item Google uses Paxos in their Chubby distributed lock service
    \item IBM supposedly uses Paxos in their SAN Volume Controller product
    \item Microsoft uses Paxos in the Autopilot cluster management service
    \item WANDisco uses Paxos in their replication technology
    \item Keyspace uses Paxos as its basic replication primitive
  \end{itemize}

@@

    \pause
    \vspace{4 mm}
    \item $B2(\beta)$ %B2 may be ensured by always picking a majority of processes as quorum, which implies that n > 2f (a majority is correct for each instance)

@@

    \pause
    \vspace{4 mm}
    \item $B3(\beta)$

@@

\pause
\vspace{3 mm}
\textsc{\textbf{Proof of Lemma}}. Let $\Psi(B, \beta)$ be defined as: $$\Psi(B, \beta) = \{B' \in \beta : (B'_{bal} > \bbalm) \wedge (B'_{val} \neq \bvalm)\}$$

@@

\pause
\vspace{3 mm}
(1) As $\Psi(B, \beta) \neq \emptyset$, we pick $C \in \Psi(B, \beta)$, such that $C_{bal} = min\{B'_{bal} : B' \in \Psi(B, \beta)\}$

\pause
\vspace{2 mm}
(2) From (1) and from the definition of $\Psi(B, \beta)$, we have that $C_{bal} > \bbalm$

@@

\pause
\vspace{2 mm}
(4) As $C_{bal} > \bbalm$ and $\bvotm \cap C_{qrm} \neq \emptyset$, we have that \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)_{bal} \geq \bbalm$

\pause
\vspace{2 mm}
(5) From (4) and the definition of \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)$, we know that such vote exists in \textit{Votes}$(\beta)$ (i.e. it is not $null$).

\pause
\vspace{2 mm}
(6) From (5) and $B3(\beta)$, we have that \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)_{val} = C_{val}$

\pause
\vspace{2 mm}
(7) From (6), and as $C_{val} \neq \bvalm$ (from $\Psi(B, \beta$)), we have that \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)_{val} \neq \bvalm$

@@

\pause
\vspace{2 mm}
(9) From (8), and since \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)_{val} \neq \bvalm$, we have that \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta) \in \text{\textit{Votes}}(\Psi(B,\beta))$

\pause
\vspace{2 mm}
(10) By the definition of \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)$, we have that \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)_{bal} < C_{bal}$

\pause
\vspace{2 mm}
(11) From (10), we have that $\exists B' : B' \in \Psi(B, \beta) \wedge B'_{bal} < C_{bal}$, which means that $$C_{bal} \neq min\{B'_{bal} : B' \in \Psi(B, \beta)\}$$ which is a contradiction with (1). $\square$

@@

\pause
\vspace{2 mm}
\textbf{\textsc{Proof of Theorem}}. If $B'_{bal} = B_{bal}$ then $B1(\beta)$ implies $B' = B$. If $B'_{bal} \neq B_{bal}$, the theorem follows immediately from the lemma. $\square$

@@

\pause
\vspace{2 mm}
\textbf{\textsc{Proof of Theorem}}.
\begin{itemize}
\item $B1(\beta \cup \{B'\})$ follows from $B1(\beta)$ and the fact that $B'_{bal} = b$, and the assumption about $b$.
\item $B2(\beta \cup \{B'\})$ follows from $B2(\beta)$ and the fact that $B'_{qrm} = Q$, and the assumption about $Q$.
\item For $B3(\beta \cup \{B'\})$, if \textit{MaxVote}$(b,Q,B) = -\infty$ then let $B'_{val}$ be any value; otherwise, let $B'_{val} = \text{\textit{MaxVote}}(b,Q,B)_{val}$. $B3(\beta \cup \{B'\})$ then follows from $B3(\beta)$. $\square$
\end{itemize} 
\end{frame}

%==========================================================
\section{The Synod Protocol}
%==========================================================

@@

\begin{itemize}
  \item To maintain $B1(\beta)$, whenever $p$ starts a new ballot $B$, it assigns a sequence number $seq$, which was never used by $p$. To avoid collisions, $p$'s unique id is appended. Then $\bbalm = (seq, p_{id})$.

  \pause
  \item To maintain $B2(\beta)$, each ballot $B$ has \bqrm\ as the majority set of the processes. As it is impossible to choose two disjoint majority sets from the same group*, $B2(\beta)$ is satisfied.

  \pause
  \item To maintain $B3(\beta)$, before initiating a ballot $B$, $p$ has to find out what is \textit{MaxVote}$(B, \bqrmm, \beta)$, which is done by finding the \textit{MaxVote}$(B, q, \beta)$ for each $q$ in \bqrm.
  \begin{itemize}
    \item If \textit{MaxVote}$(B, \bqrmm, \beta) = null$, then \bval\ can be anything;\\
    otherwise $\bvalm = \text{\textit{MaxVote}}(B, \bqrmm, \beta)_{val}$.
  \end{itemize}
\end{itemize}

\pause
{\small * If half of the processes is faulty, consistency is still respected and there is no deadlock, but the ballot $B$ does not succeed}

@@

  \pause
  \hspace{3 mm} \textbf{Phase 1a}) $p$ picks a ballot number $b$ and sends \textit{NextBallot}$(b)$ to other processes

  \pause 
  \hspace{3 mm} \textbf{Phase 1b}) each process $q$ responds with \textit{LastVote}$(b, v)$ to $p$, where $v$ is the vote cast by $q$ with the highest ballot number less than $b$ (or $null$)

@@

  \pause 
  Phase 2) Execute the ballot $B$

  \pause 
  \hspace{3 mm} \textbf{Phase 2a}) after receiving \textit{LastVote}$(b, v)$ from a majority $Q$ of the processes, $p$ initiates the ballot with number $b$ and value $val$ such that $B3(\beta)$ is maintained, by sending \textit{BeginBallot}$(b, val)$ to every process in $Q$%Bqrm = Q, Bval = val, Bbal = b

  \pause    
  \hspace{3 mm} \textbf{Phase 2b}) upon receiving \textit{BeginBallot}$(b, val)$, $q$ will not vote on it if it has already sent \textit{LastVote}$(b', v')$, such that $b' > b$; otherwise, it sends \textit{Voted}$(b, q)$ to $p$ and records this vote in its memory

@@

  \pause 
  \hspace{3 mm} \textbf{Phase 3a}) if $p$ receives \textit{Voted}$(b, q)$ from a majority $Q$ of processes, then it sends \textit{Success}$(val)$ to all processes

  \pause    
  \hspace{3 mm} \textbf{Phase 3b}) upon receiving \textit{Success}$(val)$, the process learns that the group achieved consensus on $val$.

  \pause
  \vspace{6mm}
  Every phase of this protocol is optional. Consistency is still maintained, although progress is not guaranteed.

@@

  \pause  
  Phase 2)

@@

  \pause  
  \vspace{6mm}
  Consistency is still maintained, and progress is still not guaranteed.

@@

%----------------------------------------------------------
\subsection{Complete protocol}
%----------------------------------------------------------

\begin{frame}
  \frametitle{Complete protocol}

  \begin {itemize}
    \item Same steps of the basic protocol

    \vspace{2mm}
    \item Each process performs phases 1b -- 3b as soon as they can

    \vspace{2mm}
    \item Some process must execute phase 1a
    \begin{itemize}
      \item Never initiating a ballot prevents termination
      \item However, too frequent initiation also does it %explicar cada ballot anulando o anterior, e sendo anulado antes de terminar
      \item A single process (leader) should be the only one initiating ballots %isso tb causaria a anulação de ballots anteriores de outros processos
    \end{itemize}

    \vspace{2mm}
    \item Knowledge of time, and of a maximum message delay $\Delta$, is needed
    \begin{itemize}
      \item Not exactly an asynchronous system anymore
    \end{itemize}
  \end{itemize}

@@

\begin{frame}
  \frametitle{Complete protocol}

  \begin {itemize}
    \item A minimal execution of the basic protocol would take at most $5\Delta$
%     \begin {itemize}
%       \item the responses to message (1a) would arrive after $2\Delta$ (at most)
%       \item the responses to message (2a) would arrive after $2\Delta$ (at most)
%       \item the (3a) message would take at most $\Delta$
%     \end{itemize} 
    \pause
    \vspace{2 mm}
    \item A leader would ensure progress, but that requires leader election     
    \begin {itemize}
      \item when a leader seems to have failed, a new one should be elected\\
      (the election takes time $T_{el}$)
      \item if not enough responses to (1a) or (2a) arrive after $2\Delta$, a ballot $B$ had started before $T_{el}$
      \begin{itemize}
        \item or too many failures or message losses occurred\ldots %which violates the assumptions for this protocol
      \end{itemize}
      %the other ballot may even have succeeded! 
      %maybe started by another process before T was fully elapsed
      \item in this case, a ballot with number $b' > B_{bal}$ may be started by the leader -- $NACK$s to (1a) or (2a) may have carried \bbal
      \item in the worst case, the leader receives a $NACK$ for (2a), $4\Delta$ after $T_{el}$
      \item as no other process initiated any ballot after $T_{el}$, the ballot numbered $b'$ would succeed $5\Delta$ after the $NACK$
    \end{itemize}
    \vspace{2 mm}
    \pause
    \item The complete protocol, with leader election, terminates at most after $T_{el} + 9\Delta$
    \begin{itemize}
      \item if a majority of processes didnt't fail during that time
      \item and if the leader didn't fail for $9\Delta$ after $T_{el}$
      \item and if enough messages were delivered
    \end{itemize}

  \end{itemize}

%==========================================================
\section{The Paxos Protocol}
%==========================================================

%----------------------------------------------------------
\subsection{Running multiple consensus instances}
%----------------------------------------------------------

\begin{frame}
  \frametitle{The \textbf{Paxos} Protocol}
  \framesubtitle{Multiple consensus instances}

  \begin{itemize}
    \item Each instance has an id $I_{id}$ -- and there should be a leader %although, as seen, that wouldn't be really necessary for consistency

    \pause
    \item Phases (1a) and (1b) can be executed for many of them --
    e.g. $p$ sends $NextBallot(b,n)$, for all instances $I$ with $I_{id} > n$
    \item The response from $q$ (1b) would contain a $LastVote(I_{id}, b, v)$ message for each instance $I : I_{id} > n$ that $q$ voted in
    \item Another process can send $NextBallot(b',n')$, with $b' > b$,\\
    and $p$'s (2a) messages would be ignored from then on
    \item This brings the protocol latency, on most cases, down to $3\Delta$
    \begin{itemize}
      \item no leader changes, enough correct processes and enough messages
    \end{itemize}

    \pause
    \item If each process sent messages (2b) to every other one, we'd have $2\Delta$ %show how this is done on the board

    \pause
    \item If instance $I^A$ was passed (phase (3b) concluded) before $I^B$ had a value proposed (phase (2a)), then $I^A_{id} < I^B_{id}$
    % ORDERING PROPERTY
    % the leader needed to know the result of previous instances before starting a new one (= atomic broadcast)
  \end{itemize}

%----------------------------------------------------------
\subsection{Relevance to computer science}
%----------------------------------------------------------

\begin{frame}
  \frametitle{The \textbf{Paxos} Protocol}
  \framesubtitle{Relevance to computer science}

  \begin{itemize}
    \item It may be used to implement state-machine replication
    \begin{itemize}
      \item A system (e.g. a database) may be implemented as a set of states and deterministic state changes
      \item The group of processes would be the set of replicas of the system
      \item Each state changed would be agreed upon by means of consensus (an ordered instance of Paxos)
      \item One of the replicas might be elected as leader, and client requests (state changes) would be sent to it
      \item As all replicas would have a consistent set of state changes, so would be their state
    \end{itemize} 
  \end{itemize}

  \vspace{2 mm}
  \begin{itemize}
    \item Many derivations from Paxos have been published since it was first described
    \begin{itemize}
      \item Cheap Paxos, Fast Paxos, Generalized Paxos, Multi-Paxos,\\
      Byzantine Paxos, Fast Byzantine Paxos, Fast Byzantine Multi-Paxos, Ring Paxos
    \end{itemize} 
  \end{itemize}

@@

\begin{center}
{\Huge \textbf{Thank you!}}

\vspace{5 mm}
{\footnotesize Questions?}
\end{center}
% 
%   \begin{itemize}
%     \item The paper presents the construction of a consensus protocol, based on three consistency conditions, from a preliminary one, to a complete one which ensures consistency \textbf{and} termination, as long as some message delay bound is known
%   \end{itemize}
% % 
% % The following outlook is optional.
% \vskip0pt plus.5fill
% \begin{itemize}
%   \item Outlook
%   \begin{itemize}
%     \item Something you haven't solved.
%     \item Something else you haven't solved.
%   \end{itemize}
% \end{itemize}
